---
title: "Joshua Burrows Project 1: The NHL API"
date: "Last Modified 18 September 2020"
output: 
  rmarkdown::github_document: 
    toc: TRUE
---

```{r Setup, include = FALSE}
library(assertthat)
library(knitr)
library(tidyverse)
library(rmarkdown)
library(httr)
library(jsonlite)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

This vignette explains how to access and summarize the data from the National Hockey League's APIs. 

A github repo containing the code for this vignette can be accessed using the links at the top and bottom of this page. 

The following packages were used to create this vignette: `assertthat`, `knitr`, `tidyverse`, `rmarkdown`, `httr`, and `jsonlite`. 

# Components of the getNHL() Function 

The NHL has two separate APIs: records and stats. Both can be queried using the `getNHL()` function. This function is built from a few other functions. I will begin by discussing these building blocks. 

## The Records API 

The following building block functions are used to get data from the records API. 

### Get Data 

The `getRecTable()` function grabs a table from the records API. It takes that table's extension as input. 

```{r Get Records Table}
recBaseURL <- "https://records.nhl.com/site/api"

# return the requested table from records API
getRecTable <- function(table){
  
  if(!is.string(table)){
    stop("Input must be a character string")
  }
  
  recFullURL <- paste0(recBaseURL, "/", table) 
  info <- recFullURL %>% GET() %>% content("text", encoding = "UTF-8") %>% fromJSON(flatten = "TRUE")
  info <- info[[1]] %>% as_tibble()
  
  return(info)
  
}
```

### Filter Data 

The `filterTable()` function filters records data by either franchise name or franchise id number. 

Because franchise name and franchise id number have different identifiers on different API endpoints, `filterTable()` includes `nameName` and `idNumName` arguments. In practice, the user won't have to worry about these arguments; they are behind the scenes components of `getNHL()`. 

```{r Filter Records Table}
filterTable <- function(table, nameName = NULL, idNumName = NULL, name = NULL, idNum = NULL){
  
  # if neither filter condition specified, return full table
  if((is.null(name) & is.null(idNum))){
    return(table)
  }
  
  # should search by only one condition at a time
  if((!is.null(name) & !is.null(idNum))){
    stop("search by either name or idNum, not both")
  }
  
  # either nameName or idNumName is required 
  if((!is.null(name) & is.null(nameName))){
    stop("if name is specified, nameName must be specified")
  }
  
  if((!is.null(idNum) & is.null(idNumName))){
    stop("if idNum is specified, idNumName must be specified")
  }
  
  # filter 
  if(!is.null(name)){
    if(name == "all"){ 
        return(table)
    } else{
        team <- table %>% filter(table[[nameName]] == name)
        if(length(team[[nameName]]) == 0){
          stop("not a vaild team name")
        } else{
            return(team)
          }
      } 
  }
    
  if(!is.null(idNum)){
    if(idNum == "all"){
        return(table)
    } else{
        team <- table %>% filter(table[[idNumName]] == idNum)
        if(length(team[[idNumName]]) == 0){ 
          stop("not a valid team idNum")
        } else{
            return(team)
          }
      } 
  } 
  
}
```

### Query Records Endpoints 

`getNHL()` is able to query five records endpoints. It does so by calling one of the five functions below. Each of these makes use of `getRecTable()` and `filterTable()`. Since those functions do most of the work, these five are relatively short. 

```{r Franchise Records}
# get franchise info
franchise <- function(name = NULL, idNum = NULL){

  franchises <- getRecTable("franchise") 
  
  filterTable(franchises, nameName = "teamCommonName", idNumName = "id", name, idNum) 
  
}
```

```{r Franchise Team Records}
# get franchise team totals
teamTotals <- function(name = NULL, idNum = NULL){
  
  totals <- getRecTable("franchise-team-totals") 
  
  filterTable(totals, nameName = "teamName", idNumName = "franchiseId", name, idNum) 
  
} 
```

```{r Season Records}
# get season records for a given franchise
season <- function(name = NULL, idNum = NULL){
  
  seasonRecords <- getRecTable("franchise-season-records") 
  
  filterTable(seasonRecords, nameName = "franchiseName", idNumName = "franchiseId", name, idNum) 
  
} 
```

```{r Goalie Records}
# get goalie records 
goalie <- function(name = NULL, idNum = NULL){
  
  goalieRecords <- getRecTable("franchise-goalie-records")
  
  filterTable(goalieRecords, nameName = "franchiseName", idNumName = "franchiseId", name, idNum)
  
} 
```

```{r Skater Records}
# get skater records
skater <- function(name = NULL, idNum = NULL){
  
  skaterRecords <- getRecTable("franchise-skater-records") 
  
  filterTable(skaterRecords, nameName = "franchiseName", idNumName = "franchiseId", name, idNum) 
  
}
```

## The Stats API 

Now I will turn to the building block functions that are used to contact the stats API. 

### Get Data 

`getNHL()` is able to query eight modifiers from the teams endpoint of the stats API. The job of `getStatsTable()` is to grab the data from each of the eight modified endpoints. Its length is mainly due to checking for error conditions. 

`getStatsTable()` takes the modifier as input as well as `seasonId` and `idNum`, which are additional conditions for some of the endpoints. 

```{r getStatsTable}
statsBaseURL <- "https://statsapi.web.nhl.com/api/v1/teams"

getStatsTable <- function(modifier = NULL, seasonId = NULL, idNum = NULL){
  
  # no modifier 
  if(is.null(modifier)){
    
    if((!is.null(seasonId) | !is.null(idNum))){
      stop("search by seasonId and/or idNum requires appropriate modifier")
    }
    
    base <- statsBaseURL %>% GET() %>% content("text") %>% fromJSON(flatten = TRUE)
    base <- base[[2]] %>% as_tibble()
    return(base)
  }
  
  mod <- NULL
    
  # MODIFIER 
  # team.roster 
  if(modifier == "team.roster"){
    
    if((!is.null(seasonId) | !is.null(idNum))){
      stop("search by seasonId and/or idNum not compatible with modifier team.roster")
    }
    
    mod = "?expand=team.roster"
  } 
  
  # MODIFIER 
  # person.names 
  if(modifier == "person.names"){
    
    if((!is.null(seasonId) | !is.null(idNum))){
      stop("search by seasonId and/or idNum not compatible with modifier person.names")
    }
    
    mod = "?expand=person.names" 
  }
  
  # MODIFIER 
  # team.schedule.next 
  if(modifier == "team.schedule.next"){
    
    if((!is.null(seasonId) | !is.null(idNum))){
      stop("search by seasonId and/or idNum not compatible with modifier team.schedule.next")
    }
    
    mod = "?expand=team.schedule.next"
  } 
  
  # MODIFIER 
  # team.schedule.previous 
  if(modifier == "team.schedule.previous"){
    
    if((!is.null(seasonId) | !is.null(idNum))){
      stop("search by seasonId and/or idNum not compatible with modifier team.schedule.previous")
    }
    
    mod = "?expand=team.schedule.previous"
  }
  
  # MODIFIER 
  # team.stats
  if(modifier == "team.stats"){
    
    if((!is.null(seasonId) | !is.null(idNum))){
      stop("search by seasonId and/or idNum not compatible with modifier team.stats")
    }
    
    mod = "?expand=team.stats"
  }
  
  # MODIFIER 
  # team.roster&season
  if(modifier == "team.roster&season"){
    
    if(!is.null(idNum)){
      stop("search by idNum not compatible with modifier team.roster&season")
    }
    
    if(is.null(seasonId)){
      stop("team.roster&season modifier requires search by seasonId") 
    }
    
    if(nchar(seasonId) != 8){
      stop("seasonId must be of format 20142015")
    }
    
    mod = paste0("?expand=team.roster&season=", seasonId)
  }
  
  # MODIFIER 
  # teamId
  if(modifier == "teamId"){
    
    if(!is.null(seasonId)){
      stop("search by seasonId not compatible with modifier teamId")
    }
    
    if(!is.null(idNum)){
      idCheck <- strsplit(idNum, split = "")[[1]]
      for(i in idCheck){
        if((!(i %in% c("1", "2", "3", "4", "5", "6", "7", "8", "9", "0", ",")))){
          stop("idNum must be of format 1,2,3")
        }
      }
    }
    
    mod = paste0("?teamId=", idNum)
  }
  
  # MODIFIER 
  # statsSingleSeasonPlayoffs 
  if(modifier == "statsSingleSeasonPlayoffs"){
    
    if((!is.null(seasonId) | !is.null(idNum))){
      stop("search by seasonId and/or idNum not compatible with modifier statsSingleSeasonPlayoffs")
    }
    
    mod <- "?stats=statsSingleSeasonPlayoffs"
  }
  
  if(is.null(mod)){
    stop("modifier must be one of: team.roster, person.names, team.schedule.next, team.schedule.previous, team.stats, team.roster&season, teamId, statsSingleSeasonPlayoffs")
  }
  
  # create URL and return 
  statsFullURL <- paste0(statsBaseURL, "/", mod)
  
  full <- statsFullURL %>% GET() %>% content("text") %>% fromJSON(flatten = TRUE)
  full <- full[[2]] %>% as_tibble()
  
  return(full)
  
}
```

### Flatten 

Often, the teams endpoint returns a list that contains other lists. When this happens, flattening is required. Flattening is the process of un-nesting these nested lists to return a more user friendly object. `teamStats()` does this job.

The modifier called *team.stats* returns an object with so many sub-objects that flattening it all into one 2D table creates a confusing and unwieldy data frame. So `teamStats()` doesn't try to do that. Instead, a list of data frames is returned.  

This is a long function, so check out the comments to keep your bearings. 

```{r Flatten Stats Endpoint}
teamStats <- function(modifier = NULL, seasonId = NULL, idNum = NULL){

  # get data 
  stats <- getStatsTable(modifier = modifier, seasonId = seasonId, idNum = idNum)
  
  # if no modifier is present
  if(is.null(modifier)){
    return(stats)
  }
  
  # MODIFIER 
  # team.roster
  if(modifier == "team.roster"){
    
    # get col to flatten
    roster <- stats[["roster.roster"]] 
    
    # append team names to keep track of who plays for who
    for(i in 1:length(roster)){
      roster[[i]]$name <- stats[["name"]][[i]]
    }
    
    # flatten
    roster <- do.call(rbind, roster)
    
    # join and return
    statsFlatten <- 
      left_join(roster, 
                stats, 
                by = "name", 
                suffix = c(".roster", ".base")) %>% 
      select(-c("roster.roster"))
    
    return(statsFlatten)
  } 
  
  # MODIFIER
  # person.names
  # appears to be the same as the unmodified endpoint 
  if(modifier == "person.names"){
    return(stats) 
  } 
  
  # MODIFIER
  # team.schedule.next 
  if(modifier == "team.schedule.next"){
    
    # get col to flatten
    dates <- stats[["nextGameSchedule.dates"]]
    
    # create names col to track data by team
    # create list of teams that have a next game scheduled 
    datesNonNull <- list()
    for(i in 1:length(dates)){
      if(!is.null(dates[[i]])){
        dates[[i]]$name <- stats[["name"]][[i]]
        datesNonNull <- cbind(datesNonNull, dates[i])
      }
    }
    
    # flatten datesNonNull
    for(i in 1:length(datesNonNull)){
      datesNonNull[[i]] <- 
        cbind(datesNonNull[[i]], 
              datesNonNull[[i]][["games"]]) %>% 
        select(-c("games")) 
      
      datesNonNull[[i]][["events"]] <- 
        datesNonNull[[i]][["events"]][[1]][1][[1]]
      
      datesNonNull[[i]][["matches"]] <- 
        datesNonNull[[i]][["matches"]][[1]][1][[1]]
    } 
    
    # put dates info in one tibble 
    datesVert <- tibble() 
    for(i in 1:length(datesNonNull)){
      datesVert <- rbind(datesVert, datesNonNull[[i]]) 
    } 
    
    # join flattened data with full data set 
    schedule <- 
      left_join(stats, 
                datesVert, 
                by = "name", 
                suffix = c(".main", ".nextGame")) %>% 
      select(-c("nextGameSchedule.dates"))
    
    return(schedule) 
  } 
  
  # MODIFIER
  # team.schedule.previous
  if(modifier == "team.schedule.previous"){
    
    # grab col to flatten
    dates <- stats[["previousGameSchedule.dates"]]
    
    # create names column to track data by team
    # create list of teams that had a previous game 
    datesNonNull <- list()
    for(i in 1:length(dates)){
      if(!is.null(dates[[i]])){
        dates[[i]]$name <- stats[["name"]][[i]]
        datesNonNull <- cbind(datesNonNull, dates[i])
      }
    }
    
    # flatten datesNonNull
    for(i in 1:length(datesNonNull)){
      datesNonNull[[i]] <- 
        cbind(datesNonNull[[i]], 
              datesNonNull[[i]][["games"]]) %>% 
        select(-c("games")) 
      
      datesNonNull[[i]][["events"]] <- 
        datesNonNull[[i]][["events"]][[1]][1][[1]]
      
      datesNonNull[[i]][["matches"]] <- 
        datesNonNull[[i]][["matches"]][[1]][1][[1]]
    } 
    
    # create venue.id col where it doesn't exist 
    for(i in 1:length(datesNonNull)){
      if(is.null(datesNonNull[[i]][["venue.id"]])){
        datesNonNull[[i]][["venue.id"]] <- NA
      }
    }
    
    # put dates info in one tibble 
    datesVert <- tibble() 
    for(i in 1:length(datesNonNull)){
      datesVert <- rbind(datesVert, datesNonNull[[i]]) 
    } 
  
    # join flattened data with full data set 
    schedule <- 
      left_join(stats, 
                datesVert, 
                by = "name", 
                suffix = c(".main", ".previousGame")) %>% 
      select(-c("previousGameSchedule.dates"))
    
    return(schedule) 
  } 
  
  # MODIFIER
  # team.stats
  if(modifier == "team.stats"){ 
    
    # create empty list to contain final results
    final <- list()
    
    # include data from the base endpoint 
    final$teamOverview <- stats %>% select(-c("teamStats"))
    
    # get col to flatten
    teamStats <- stats[["teamStats"]]
    
    # empty tibbles to contain flattened data 
    statsOverview <- tibble()
    statsValues <- tibble()
    statsRanks <- tibble()
    
    # flatten 
    for(i in 1:length(teamStats)){
      teamStats[[i]]$name <- stats[["name"]][[i]]
      
      statsOverview <- rbind(statsOverview, teamStats[[i]])
      
      statsValues <- rbind(statsValues, teamStats[[i]][["splits"]][[1]][1, ]) 
      
      statsRanks <- rbind(statsRanks, teamStats[[i]][["splits"]][[1]][2, ])
    }
    
    # values table should be numeric 
    for(i in 1:(ncol(statsValues) - 2)){
      statsValues[[i]] <- as.double(statsValues[[i]])
    }
    
    # fill in final list 
    statsOverview <- statsOverview %>% select(-c("splits"))
    final$statsOverview <- statsOverview
    final$statsValues <- statsValues
    final$statsRanks <- statsRanks
    
    return(final)
  } 
  
  # MODIFIER
  # team.roster&season
  if(modifier == "team.roster&season"){
    
    # get col to flatten 
    roster <- stats[["roster.roster"]]
    
    # add team names to keep track of who plays for who 
    for(i in 1:length(roster)){
      roster[[i]]$name <- stats[["name"]][[i]]
    }
    
    # flatten 
    roster <- do.call(rbind, roster)
    
    # join and return 
    statsFlatten <- 
      left_join(roster, 
                stats, 
                by = "name", 
                suffix = c(".roster", ".base")) %>% 
      select(-c("roster.roster"))
    
    return(statsFlatten)
  }
  
  # MODIFIER
  # teamId
  # getStatsTable() does all that is needed here 
  if(modifier == "teamId"){
    return(stats) 
  }
  
  # MODIFIER
  # statsSingleSeasonPlayoffs
  # appears to be the same as the unmodified endpoint
  if(modifier == "statsSingleSeasonPlayoffs"){
    return(stats)
  }
  
}
```

# Wrapper Function: getNHL()

Until now, we have been looking at functions that live behind the scenes. The star of the show is `getNHL()`. 

`getNHL()` allows the user to query both the records API and the stats API. It is a wrapper function that works by calling other functions. 

`getNHL()` has six arguments, but you never use all six at once: 

* API: "records"  
  - recTable: the desired records endpoint  
  - name: filter by franchise name  
  - idNum: filter by franchise id  
* API: "stats"  
  - modifier: the desired modifier  
  - idNum: for the *teamId* modifier  
  - seasonId: for the *team.roster&season* modifier  

```{r getNHL() Wrapper}
getNHL <- function(API, recTable = NULL, name = NULL, idNum = NULL, modifier = NULL, seasonId = NULL){
  
  # check error conditions 
  if(!(API %in% c("records", "stats"))){
    stop("API must be either records or stats")
  }
  
  if((API == "records" & is.null(recTable))){
    stop("recTable must be specified to query records API")
  }
  
  if((API == "records" & (!is.null(seasonId) | !is.null(modifier)))){
    stop("seasonId and modifier are not compatible with records API")
  }
  
  if((API == "stats" & (!is.null(recTable) | !is.null(name)))){
    stop("recTable and name are not compatible with stats API")
  }
  
  # query records API
  if(API == "records"){
    
    if(recTable == "franchise"){
      tab <- franchise(name = name, idNum = idNum)
      return(tab)
    }
    
    if(recTable == "franchise-team-totals"){
      tab <- teamTotals(name = name, idNum = idNum)
      return(tab)
    }
    
    if(recTable == "franchise-season-records"){
      tab <- season(name = name, idNum = idNum)
      return(tab) 
    }
    
    if(recTable == "franchise-goalie-records"){
      tab <- goalie(name = name, idNum = idNum)
      return(tab) 
    }
    
    if(recTable == "franchise-skater-records"){
      tab <- skater(name = name, idNum = idNum)
      return(tab) 
    }
    
    stop("recTable must be one of the following strings: franchise, franchise-team-totals, franchise-season-records, franchise-goalie-records, franchise-skater-records")
  
  }
  
  # query stats API 
  if(API == "stats"){
   tab <- teamStats(modifier = modifier, seasonId = seasonId, idNum = idNum)
   return(tab)
  }
  
}
```

## Examples 

Below are a few queries that demonstrate how `getNHL()` is used. 

```{r getNHL() Examples}
getNHL(API = "records", recTable = "franchise") %>% head()

getNHL(API = "records", recTable = "franchise-team-totals", name = "Carolina Hurricanes")

getNHL(API = "records", recTable = "franchise-season-records", idNum = 15) 

getNHL(API = "stats", modifier = "team.schedule.previous") %>% head()

getNHL(API = "stats", modifier = "teamId", idNum = "1,5,10")

```

# Exploratory Data Analysis 

Now that we can get data about the NHL, let's take a look at it! 

## Wins and Penalty Minutes 

### Regular Season

Unsurprisingly, there is a negative correlation between wins and penalty minutes during the regular season. 

It is worth noting, though, that many teams with high win percentages still have substantial penalty minutes. Perhaps this is because players foul strategically; they get something in exchange for minutes spent in the box. 

Teams that have played less than 500 regular season games were not considered. 

```{r Wins and Penalties Regular}
totals <- getNHL(API = "records", recTable = "franchise-team-totals")
totalsRegular <- totals %>% filter(gameTypeId == 2)

totalsRegular <- totalsRegular %>% 
  filter(gamesPlayed >= 500) %>% 
  mutate(winPct = wins / gamesPlayed, 
         penaltyAvg = penaltyMinutes / gamesPlayed)

corRegular <- cor(totalsRegular$winPct, totalsRegular$penaltyAvg)

g <- ggplot(totalsRegular, aes(x = penaltyAvg, y = winPct)) 

g + 
  geom_point() + 
  geom_smooth(method = lm, col = "red") + 
  geom_text(x = 20, y = .475, label = paste0("Correlation = ", round(corRegular, 2))) + 
  labs(title = "Wins and Penalty Minutes, Regular Season", 
       x = "Average Penalty Minutes", 
       y = "Win Percentage", 
       caption = "Teams with at least 500 games")
```

### Post Season 

During the post season, the negative correlation between win percentage and penalty minutes is not as strong. Perhaps this is because, during the post season, players are even better at fouling strategically. 

Teams with less than 50 post season games were not considered. 

```{r Wins and Penalties Post}
totalsPost <- totals %>% filter(gameTypeId == 3)

totalsPost <- 
  totalsPost %>% 
  filter(gamesPlayed >= 50) %>% 
  mutate(winPct = wins / gamesPlayed, 
         penaltyAvg = penaltyMinutes / gamesPlayed)

corPost <- cor(totalsPost$winPct, totalsPost$penaltyAvg)

g2 <- ggplot(totalsPost, aes(x = penaltyAvg, y = winPct)) 

g2 + 
  geom_point() + 
  geom_smooth(method = lm, col = "red") + 
  geom_text(x = 25, y = .55, label = paste0("Correlation = ", round(corPost, 2))) + 
  labs(title = "Wins and Penalty Minutes, Post Season", 
       x = "Average Penalty Minutes", 
       y = "Win Percentage", 
       caption = "Teams with at least 50 games")
```

## Win/Loss Differential and Shots Differential 

### Join and Calculate 

Below, a new table is created by joining data from the *team.stats* modifier with data from the unmodified stats endpoint. Then three new variables are calculated: 

* Win/Loss Differential: Wins - Losses  
* Win/Loss Diff Indicator: Whether a team has a winning record  
* Shots Differential: Shots taken - Shots allowed  

```{r Diffs1}
teamInfo <- getNHL(API = "stats", modifier = "team.stats")
values <- teamInfo[["statsValues"]]

teams <- getNHL(API = "stats")

teamValues <- 
  inner_join(teams, 
             values, 
             by = c("name" = "team.name"), 
             suffix = c(".teams", ".values"))

teamValues <- 
  teamValues %>% 
  mutate(WLDiff = stat.wins - stat.losses, 
         shotDiff = stat.shotsPerGame - stat.shotsAllowed)

positive <- function(x){
  if(x > 0){
    return("Winning Team")
  } else if(x == 0){
    return("50/50 Team")
    } else{
        return("Losing Team")
      }
}

positive <- Vectorize(positive)

teamValues <- teamValues %>% mutate(WLDiffIndicator = positive(WLDiff))

teamValues %>% select(c("name", "WLDiff", "shotDiff", "WLDiffIndicator")) %>% head()
```

### Win/Loss Differential by Division and Conference 

Every team in the **Central** division is a winning team. In contrast, the **Atlantic** division has the worst winning team to losing team ratio. 

The **Western** conference has a better winning team to losing team ratio than the **Eastern** conference. 

Overall, the NHL has more winning teams than losing teams.

This data is visualized in contingency tables and bar charts.

```{r Diffs2}
table(teamValues$division.name, teamValues$WLDiffIndicator) %>% kable(caption = "Divisions")

table(teamValues$conference.name, teamValues$WLDiffIndicator) %>% kable(caption = "Conferences")

g3 <- ggplot(teamValues, aes(x = WLDiffIndicator))

g3 + 
  geom_bar(aes(fill = division.name), position = "dodge") + 
  labs(title = "Winning Teams by Division", 
       x = "Record", 
       y = "Count") + 
  scale_fill_discrete(name = "Division")

g3 + 
  geom_bar(aes(fill = conference.name), position = "dodge") + 
  labs(title = "Winning Teams by Conference", 
       x = "Record", 
       y = "Count") + 
  scale_fill_discrete(name = "Conference")
```

### Win/Loss Differential Histogram 

The predominance of winning teams in the NHL can be visualized with a histogram. The density line peaks at a win/loss differential of 10. This indicates win/loss differentials in the vicinity of 10 are fairly common. 

```{r Diffs3} 
g4 <- ggplot(teamValues, aes(x = WLDiff, y = ..density..))

g4 + 
  geom_histogram(bins = 10, fill = "lightblue") + 
  geom_density(lwd = 2, color = "darkblue") + 
  labs(title = "Win/Loss Differential", 
       x = "Wins Minus Losses", 
       y = "Density")
```

### Shots Differential Vs Win/Loss Differential 

Unsurprisingly, there is a positive correlation between win/loss differential and shots differential. The slope of the regression line is steeper in the **Eastern** conference than the **Western** conference. This slope is not changed much by removing an outlier. 

```{r Diffs4}
g5 <- ggplot(teamValues, aes(x = WLDiff, y = shotDiff, color = conference.name)) 

g5 + 
  geom_point() + 
  geom_smooth(method = lm) + 
  labs(title = "Shots Taken/Allowed Differential vs Win/Loss Differential", 
       x = "Wins Minus Losses", 
       y = "Shots Taken Minus Shots Allowed", 
       color = "Conference", 
       caption = "With Outlier") +
  coord_cartesian(xlim = c(-31, 31), ylim = c(-8, 8))

teamValuesNoOut <- teamValues %>% filter(WLDiff > -20)

g6 <- ggplot(teamValuesNoOut, aes(x = WLDiff, y = shotDiff, color = conference.name)) 

g6 + 
  geom_point() + 
  geom_smooth(method = lm) + 
  labs(title = "Shots Taken/Allowed Differential vs Win/Loss Differential ", 
       x = "Wins Minus Losses", 
       y = "Shots Taken Minus Shots Allowed", 
       color = "Conference", 
       caption = "Without Outlier") +
  coord_cartesian(xlim = c(-31, 31), ylim = c(-8, 8))
```

## Goals Per Game 

### Calculate and Join 

Below, two new variables are calculated from the regular season team totals data: 

* Avg Goals Against: goals against divided by total games  
* Avg Goals For: goals for divided by total games  

The regular season team totals data is then joined with data from the base endpoint of the stats API. The result is a table with information about active teams. Again, teams with less than 500 regular season games were not considered. 

```{r Active Team Totals}
totalsRegular <- 
  totalsRegular %>% 
  mutate(avgGoalsAgainst = goalsAgainst / gamesPlayed, 
         avgGoalsFor = goalsFor / gamesPlayed)

totalsRegularActive <- 
  left_join(teams, 
            totalsRegular, 
            by = c("name" = "teamName"), 
            suffix = c(".teams", ".stats"))

totalsRegularActive %>% select(c("name", "avgGoalsAgainst", "avgGoalsFor")) %>% head()
```

### Regular Season Goals by Division 

The **Central** division has the lowest average goals for and the lowest average goals against. The **Pacific** division has the highest average goals for and the highest average goals against. 

The boxplots below show significant variation between teams within a division. This is emphasized by the superimposed scatter points. 

All numbers are for the regular season. 

```{r Avg Goals}
avgs <- totalsRegularActive %>% 
  group_by(division.name) %>% 
  summarize(divisionAvgGoalsAgainst = round(mean(avgGoalsAgainst, na.rm = TRUE), 3), 
            divisionAvgGoalsFor = round(mean(avgGoalsFor, na.rm = TRUE), 3))

avgs %>% 
  kable(col.names = c("Division", 
                      "Divison Avg Goals Against", 
                      "Division Avg Goals For"), 
        caption = "Teams with at least 500 games")

g7 <- ggplot(totalsRegularActive, aes(x = division.name))

g7 + 
  geom_boxplot(aes(y = avgGoalsAgainst)) + 
  geom_point(aes( y = avgGoalsAgainst, 
                  color = division.name), 
             position = "jitter") + 
  labs(title = "Average Goals Against by Division", 
       x = "Division", 
       y = "Average Goals Against", 
       color = "Division", 
       caption = "Teams with at least 500 games") 

g7 + 
  geom_boxplot(aes(y = avgGoalsFor)) + 
  geom_point(aes(y = avgGoalsFor, 
                 color = division.name), 
             position = "jitter") + 
  labs(title = "Average Goals For by Division", 
       x = "Division", 
       y = "Average Goals For", 
       color = "Division", 
       caption = "Teams with at least 500 games")
```

## Regular Season Numeric Summaries by Division

Six number summaries are calculated for three variables for for active teams in each division. All numbers are for the regular season. Again, teams with less than 500 games were not considered. 

One thing that jumps out about these summaries is that average goals for and average goals against are so similar across divisions. This suggests that no one division is far and away better than the others. This is confirmed by the similarity in point percentage across divisions. 

```{r Sums by Divison 1}
# function that takes in division name and returns summary table
divSum <- function(div){
  
  # make sure input is correct type
  if(!is.string(div)){
    stop("Input must be a character string!") 
  } 
  
  # get the numbers for the division in question
  splitz <- 
    totalsRegularActive %>% 
    filter(division.name == div) %>% 
    select(c("pointPctg", 
             "avgGoalsAgainst", 
             "avgGoalsFor")) %>% 
    drop_na()
  
  # get the summary statistics 
  statz <- 
    splitz %>% 
    apply(MARGIN = 2, FUN = summary) %>% 
    round(digits = 3)
  
  # format pretty for printing 
  cap <- paste0(div, " Division")
  
  statz <- statz %>% 
    kable(caption = cap, 
          col.names = c("Point Percentage",
                        "Average Goals Against", 
                        "Average Goals For"))
  
  return(statz)
  
}
```

```{r Sums by Division 2}
divSum("Atlantic")
divSum("Central")
divSum("Metropolitan")
divSum("Pacific")
```





